Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-13 paper/comoving-followup.tex
columns of data tables

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}

@@

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{sample_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
    blue histogram) and all comoving star pair members observed in this
    follow-up effort (black points); colored lines connect candidate comoving
    star pairs.
    2MASS $J$-band magnitudes are obtained from the Gaia science
    archive.
    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
    \citealt{Oh:2017}).
    \label{fig:sample-cmd}}
\end{figure}

@@

For each source spectrum, with the resulting posterior samples, we compute the
median and median absolute deviation (MAD) of the marginal distribution over
line centroid, $x_0$.
We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
MAD}$, assuming the samples are Gaussian distributed) of the samples and find
that the typical velocity \emph{precision} from measuring the centroid of \Ha\
is between $5$--$10~\kms$.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
  \end{center}
  \caption{%
    A corner plot showing all projections of posterior PDF samples for the \Ha\
    line model and the spectrum of the source HD 63408.
    Note the small uncertainty in the line centroid measurement, $x_0$, which
    corresponds to a $\approx 2~\kms$ velocity precision for this source.
    Parameters are described in \sectionname~\ref{sec:reduction}, and prior PDFs
    over the parameters are described in \tablename~\ref{tbl:prior-bounds}.
    \label{fig:Halpha-mcmc-corner}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{mcmc_example_fit.pdf}
  \end{center}
  \caption{%
    Visualization of the \Ha\ absorption line model and fits for the source HD
    63408 as a demonstration of the procedure used for all targets.
    \emph{Top}:
    The source spectrum in a small wavelength range around \Ha\ (black),
    uncertainty in spectral flux (gray error bars).
    Over-plotted is the line model + background computed from the median
    posterior PDF sample (dark green line) and the 15th to 85th percentile
    region computed from all posterior PDF samples (shaded green region).
    \emph{Bottom}:
    Source spectrum residuals computed by subtracting the median model (dark
    green line in top panel) from the source spectrum.
    The over-plotted line (orange) shows the Gaussian process background model
    computed from the median posterior PDF sample.
    \label{fig:Halpha-mcmc-fit}}
\end{figure}

@@

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of confirmed comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs} and
    \figurename~\ref{fig:highlighted-cmd}).
    \label{fig:genuine-cmd}}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{highlighted_cmd.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:genuine-cmd}, but only three highlighted
    comoving star pairs are shown.
    \label{fig:highlighted-cmd}}
\end{figure}

We compile Tycho 2 ($B$, $V$; \citealt{derp}), 2MASS ($J$, $H$, $K_{\rm s}$;
\citealt{derp}), and WISE ($W_1$, $W_2$, $W_3$; \citealt{derp}) photometry for
each star in the pairs below and fit isochrones to the stars individually to
measure their stellar parameters.
We use the \package{isochrones} \python\ package (\citealt{Morton:2015}) to fit
the photometric and parallax information (and uncertainties) for each star to
infer the \logg, \teff, and stellar mass; these values are reported in
\tablename~\ref{tbl:highlighted-pairs}.
In brief, we use the model grids from the Dartmouth Stellar Evolution Database
(\citealt{Dotter:2008}) and the built-in interpolation in \package{isochrones}
to compute the posterior probability of a given set of stellar parameters given
the data (photometry and parallax).
We generate samples (in stellar parameters) from this posterior pdf using the
ensemble MCMC sampler implemented in \package{emcee}
(\citealt{Foreman-Mackey:2013}) using 128 walkers.
We run the sampler for 256 initial steps from random initial positions, compute
the median position of the walkers in parameter-space and re-initialize the
walkers in a small Gaussian ball around this median position to run for another
1024 burn-in steps.
We then discard all but the last position and re-run for a final 8192 steps,
saving only every 128th step.
The stellar parameter values listed below are computed by taking the median
posterior sample, and the uncertainties are estimated by computing the median
absolute deviation, ${\rm MAD}$, and converting to a standard deviation under
the assumption of Gaussianity ($\sigma \approx 1.5 \times {\rm MAD}$).

@@

@@

Comoving stars are also sensitive potentiometers of the Galactic mass
distribution.
- Long-term goal is to study the separation distribution of wide-binaries as a
  function of age and Galactic orbit distribution (disk vs. halo)
    - Is there a separation truncation?
    - Can that be described just from Galactic tides, or need perturbers?

% Caution that prior results are for red giants, MS stars may have planetary
% imprints that get erased as dredge-up and convective layer erases...

@@

Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>

@@

% Low-resolution spectroscopic follow-up of comoving pairs.

% Using relative RV's, confirm a sample of very wide separation comoving star pairs.

% Either wide binaries, unbound binaries, or dissolving clusters

% Need much larger sample with no cuts on separation or signal-to-noise to study
% true separation distribution

>>>>

% Project-specific
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\DR}[1]{\acronym{DR}#1}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\tmass}{\acronym{2MASS}}
\newcommand{\lbfgsb}{\texttt{L-BFGS-B}}
\newcommand{\Ha}{\ensuremath{{\rm H}\alpha}}
\newcommand{\dex}{\ensuremath{{\rm dex}}}
\newcommand{\logg}{\ensuremath{\ensuremath{\log g}}}
\newcommand{\teff}{\ensuremath{\ensuremath{T_{\rm eff}}}}
\newcommand{\llrold}{\ensuremath{\mathcal{R}_\mu}}
\newcommand{\llrnew}{\ensuremath{\mathcal{R}_{\rm RV}}}

@@

\begin{figure}[ht!]
  \begin{center}
    \includegraphics[width=0.8\linewidth]{sample_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram for all \tgas\ sources within $200~\pc$ (background
    blue histogram) and all comoving star pair members observed in this
    follow-up effort (black points); colored lines connect candidate comoving
    star pairs.
    \tmass\ (\citealt{Skrutskie:2006}) $J$-band magnitudes are obtained
    from the Gaia science archive.
    Apparent $G$-band magnitudes (\citealt{Carrasco:2016}) are converted to
    absolute magnitude, $M_G$, using a Lutz-Kelker-corrected
    (\citealt{Lutz:1973}) distance estimate (see also \eqname~1 in
    \citealt{Oh:2017}).
    \label{fig:sample-cmd}}
\end{figure}

@@

@@

For each source spectrum, with the resulting posterior samples, we compute the
median and median absolute deviation (MAD) of the marginal distribution over
line centroid, $x_0$.
We use the MAD to estimate the standard deviation ($\sigma \approx 1.5\times{\rm
MAD}$, assuming the samples are Gaussian distributed) of the samples and find
that the typical velocity \emph{precision} from measuring the centroid of \Ha\
is between $5$--$10~\kms$.
We report the measured absolute radial velocities along with compiled sky
position, astrometry, and $G$-band magnitude from the \tgas\ catalog and \tmass\
$J$-band magnitudes for all of our comoving star targets in the ``stars'' data
table available with this \documentname; see \tablename~\ref{tbl:data-stars} for
a detailed description of the content of this table.

\begin{figure}[!hp]
  \begin{center}
    \includegraphics[width=\linewidth]{mcmc_corner.pdf}
  \end{center}
  \caption{%
    A corner plot showing all projections of posterior PDF samples for the \Ha\
    line model and the spectrum of the source HD 63408.
    Note the small uncertainty in the line centroid measurement, $x_0$, which
    corresponds to a $\approx 2~\kms$ velocity precision for this source.
    Parameters are described in \sectionname~\ref{sec:reduction}, and prior PDFs
    over the parameters are described in \tablename~\ref{tbl:prior-bounds}.
    \label{fig:Halpha-mcmc-corner}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.7\linewidth]{mcmc_example_fit.pdf}
  \end{center}
  \caption{%
    Visualization of the \Ha\ absorption line model and fits for the source HD
    63408 as a demonstration of the procedure used for all targets.
    \emph{Top}:
    The source spectrum in a small wavelength range around \Ha\ (black),
    uncertainty in spectral flux (gray error bars).
    Over-plotted is the line model + background computed from the median
    posterior PDF sample (dark green line) and the 15th to 85th percentile
    region computed from all posterior PDF samples (shaded green region).
    \emph{Bottom}:
    Source spectrum residuals computed by subtracting the median model (dark
    green line in top panel) from the source spectrum.
    The over-plotted line (orange) shows the Gaussian process background model
    computed from the median posterior PDF sample.
    \label{fig:Halpha-mcmc-fit}}
\end{figure}

\begin{table}[htb]
    \centering
    \caption{Description of columns in data table containing results from radial
    velocity follow-up.} \label{tbl:data-stars}
    \begin{tabular}{l|l|l}
        \toprule
        Column Name    & Unit & Description\\
        \midrule
        \texttt{Oh17\_id}         &      & Row index in candidate catalog (\citealt{Oh:2017})\\
        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
        \texttt{tgas\_source\_id} &      & Unique source identifier from \tgas\\
        \texttt{name}             &      & HD, \project{Hipparcos}, or \project{Tycho-2} identifier\\
        \texttt{ra}               & deg  & Right ascension from \tgas\\
        \texttt{dec}              & deg  & Declination from \tgas\\
        \texttt{parallax}         & mas  & Parallax from \tgas\\
        \texttt{distance}         & pc   & Distance estimate using inverse parallax from \tgas\\
        \texttt{G}                & mag  & \gaia\ $G$-band magnitudes\\
        \texttt{J}                & mag  & \tmass\ $J$-band magnitudes\\
        \texttt{rv}               & \kms & Radial velocity measured in this work\\
        \texttt{rv\_err}          & \kms & Radial velocity error\\
        \bottomrule
    \end{tabular}
\end{table}

% \clearpage

@@

The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
all observed candidate comoving pairs are provided with the associated ``pairs''
table; see \tablename~\ref{tbl:data-pairs} for a detailed description of the
content of this table.

\begin{table}[htb]
    \centering
    \caption{Description of columns in data table containing results from radial
    velocity follow-up.} \label{tbl:data-stars}
    \begin{tabular}{l|l|l}
        \toprule
        Column Name    & Unit & Description\\
        \midrule
        \texttt{Oh17\_group\_id}  &      & Group ID from candidate catalog (\citealt{Oh:2017})\\
        \texttt{R\_mu}            &      & Log-likelihood-ratio computed with only astrometric information\\
        \texttt{R\_rv}            &      & Log-likelihood-ratio computed with radial velocities as well\\
        \texttt{dx\_50}           & \pc  & Median of 3D spatial separation error distribution\\
        \texttt{dv\_50}           & \kms & Median of 3D velocity difference error distribution\\
        \texttt{dx\_15}           & \pc  & 15th percentile of 3D spatial separation error distribution\\
        \texttt{dx\_85}           & \pc  & 85th percentile of 3D spatial separation error distribution\\
        \texttt{dv\_15}           & \kms & 15th percentile of 3D velocity difference error distribution\\
        \texttt{dv\_85}           & \kms & 85th percentile of 3D velocity difference error distribution\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.8\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of confirmed comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the confirmed comoving pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs} and
    \figurename~\ref{fig:highlighted-cmd}).
    \label{fig:genuine-cmd}}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.8\linewidth]{highlighted_cmd.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:genuine-cmd}, but only three highlighted
    comoving star pairs are shown.
    \label{fig:highlighted-cmd}}
\end{figure}

We compile \project{Tycho-2} ($B$, $V$; \citealt{Hog:2000}), \tmass\
($J$, $H$, $K_{\rm s}$; \citealt{Skrutskie:2006}), and \acronym{WISE} ($W_1$,
$W_2$, $W_3$; \citealt{Wright:2010}) photometry for each star in the pairs below
and fit isochrones to the stars individually to measure their stellar
parameters.
We use the \package{isochrones} \python\ package (\citealt{Morton:2015}) to fit
the photometric and parallax information (and uncertainties) for each star to
infer the \logg, \teff, and stellar mass; these values are reported in
\tablename~\ref{tbl:highlighted-pairs}.
In brief, we use the model grids from the Dartmouth Stellar Evolution Database
(\citealt{Dotter:2008}) and the built-in interpolation in \package{isochrones}
to compute the posterior probability of a given set of stellar parameters given
the data (photometry and parallax).
We generate samples (in stellar parameters) from this posterior pdf using the
ensemble MCMC sampler implemented in \package{emcee}
(\citealt{Foreman-Mackey:2013}) using 128 walkers.
We run the sampler for 256 initial steps from random initial positions, compute
the median position of the walkers in parameter-space and re-initialize the
walkers in a small Gaussian ball around this median position to run for another
1024 burn-in steps.
We then discard all but the last position and re-run for a final 8192 steps,
saving only every 128th step.
The stellar parameter values listed below are computed by taking the median
posterior sample, and the uncertainties are estimated by computing the median
absolute deviation, ${\rm MAD}$, and converting to a standard deviation under
the assumption of Gaussianity ($\sigma \approx 1.5 \times {\rm MAD}$).

@@

\subsection{Caveats} \label{sec:caveats}

Need much larger sample with no cuts on separation or signal-to-noise to study
true separation distribution

@@

@@

% \subsection{Prospects for chemical tagging} \label{sec:chemical-tagging}

@@

% \subsection{Prospects for Galactic structure} \label{sec:dynamics}

\subsection{Future spectroscopic surveys} \label{sec:future-rvs}

DISCO / AS4, 4MOST, WEAVE, DESI, PFS
